---
title: "SILAC H/L Ratios for Histone PTMs"
author: "David Shechter"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
---

This R Markdown document analyzes SILAC-labeled histone post-translational modifications (PTMs) in response to PRMT5 inhibition using GSK591. Quantitative peptide intensity data were obtained for histone H3 and H4 from nascent proteome experiments and processed to compute three key metrics:

1. **Fraction Heavy (H / (H + L))** – reflecting newly synthesized peptide incorporation,
2. **PTM Occupancy** – relative abundance of each PTM site within a peptide,
3. **Normalized Peptide Abundance** – median-centered abundance across replicates.

Data are imported from Excel files, reshaped to long format, and parsed by treatment condition and timepoint. The script generates:

- Summary statistics per PTM and condition
- QC plots for each metric
- Heatmaps of fraction heavy values per peptide
- Stacked barplots separating old vs. new histone PTMs

Exported outputs include per-peptide and summary-level data tables (in `.xlsx` and `.csv`), and all plots (in `.pdf`) are saved to the output directory.

This workflow enables visualization and interpretation of histone PTM dynamics under control (DMSO) and PRMT5-inhibited conditions at 9 hr and 24 hr timepoints.


```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(writexl)
library(pheatmap)   # for heatmaps
library(stringr)
library(glue)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(circlize)
library(RColorBrewer)
```

```{r load and prepare data}
# Load Excel sheets
path <- "M:/Github_local/Roth_etal_2025/Figure 6/SILAC/SILAC_HistonePTM_May2025.xlsx"

outdir <- "M:/Github_local/Roth_etal_2025/Figure 6/SILAC/"

dir.create(paste0(outdir,"plots"), showWarnings = FALSE)
dir.create(paste0(outdir, "csv"), showWarnings = FALSE)



df_H3 <- read_excel(path, sheet="simplified_H3") %>% mutate(Histone = "H3")
df_H4 <- read_excel(path, sheet="simplified_H4") %>% mutate(Histone = "H4")
int_df <- bind_rows(df_H3, df_H4)

df_H3_abund <- read_excel(path, sheet="Abundance_H3") %>% mutate(Histone = "H3")
df_H4_abund <- read_excel(path, sheet="Abundance_H4") %>% mutate(Histone = "H4")
abund_df <- bind_rows(df_H3_abund, df_H4_abund)

intensity_cols <- names(int_df)[str_detect(names(int_df), '^[0-9]+_[DG]-[0-9]{1,2}hr_rep\\d+$')]
abund_cols     <- names(abund_df)[str_detect(names(abund_df), '^[0-9]+_[DG]-[0-9]{1,2}hr_rep\\d+$')]

# Pivot to long format and parse sample info
int_long <- int_df %>%
  pivot_longer(all_of(intensity_cols), names_to="sample", values_to="Intensity") %>%
  mutate(sample = str_remove(sample, '^[0-9]+_')) %>%
  separate(sample, into=c("Treatment","Time","Rep"), sep="[-_]", remove=FALSE) %>%
  mutate(
    Treatment = recode(Treatment, D="DMSO", G="GSK591"),
    Time      = str_replace(Time, "hr", " hr"),
    Rep       = parse_number(Rep)
  )

abundance_long <- abund_df %>%
  pivot_longer(all_of(abund_cols), names_to="sample", values_to="Abundance") %>%
  mutate(sample = str_remove(sample, '^[0-9]+_')) %>%
  separate(sample, into=c("Treatment","Time","Rep"), sep="[-_]", remove=FALSE) %>%
  mutate(
    Treatment = recode(Treatment, D="DMSO", G="GSK591"),
    Time      = str_replace(Time, "hr", " hr"),
    Rep       = parse_number(Rep)
  )


```

```{r compute-and-export-fraction, message=FALSE}

# COMPUTE FRACTION HEAVY AND EXPORT
# Compute Fraction Heavy
int_sum <- int_long %>%
  group_by(Peptide_sequence, PTM, Histone, Treatment, Time, Rep, Heavy_Light) %>%
  summarize(total_int = sum(Intensity, na.rm=TRUE), .groups="drop")

frac_df <- int_sum %>%
  pivot_wider(
    id_cols = c(Peptide_sequence, PTM, Histone, Treatment, Time, Rep),
    names_from = Heavy_Light,
    values_from = total_int,
    values_fill = 0
  ) %>%
  mutate(fracHeavy = heavy / (heavy + light))

summary_fracH <- frac_df %>%
  group_by(Peptide_sequence, PTM, Histone, Treatment, Time) %>%
  summarize(
    mean_fracH = mean(fracHeavy, na.rm=TRUE),
    sd_fracH   = sd(fracHeavy,   na.rm=TRUE),
    .groups="drop"
  ) %>%
  mutate(RowID = glue("{Peptide_sequence} | {PTM}"))

# Compute normalized occupancy per peptide
abund_occ <- abundance_long %>%
  group_by(Peptide_sequence, PTM, Histone, Treatment, Time, Rep) %>%
  mutate(
    total_pep = sum(Abundance, na.rm=TRUE),
    occ = Abundance / total_pep
  ) %>%
  ungroup()

summary_occ <- abund_occ %>%
  group_by(Peptide_sequence, PTM, Histone, Treatment, Time) %>%
  summarize(
    mean_occ = mean(occ, na.rm=TRUE),
    sd_occ   = sd(occ,   na.rm=TRUE),
    .groups="drop"
  ) %>%
  mutate(RowID = glue("{Peptide_sequence} | {PTM}"))

abundance_norm <- abundance_long %>%
  group_by(Treatment, Time, Rep, Histone) %>%
  mutate(
    sample_median = median(Abundance, na.rm = TRUE),
    norm_abund    = Abundance / sample_median
  ) %>%
  ungroup()

# Abundance (raw)
summary_abund <- abundance_norm %>%
  group_by(Peptide_sequence, PTM, Histone, Treatment, Time) %>%
  summarize(
    mean_abund = mean(norm_abund, na.rm=TRUE),
    sd_abund   = sd(norm_abund,   na.rm=TRUE),
    .groups="drop"
  ) %>%
  mutate(RowID = glue("{Peptide_sequence} | {PTM}"))

######### EXPORT

write_xlsx(
  list(
    per_peptide = frac_df,
    summary     = summary_fracH
  ),
  path = "HistonePTM_fracHeavy_PRMT5.xlsx"
)

write_xlsx(
  list(
    per_peptide = abund_occ,
    summary     = summary_occ
  ),
  path = "HistonePTM_Occupancy_PRMT5.xlsx"
)

write_xlsx(
  list(
    per_peptide = abundance_long,
    summary     = summary_abund
  ),
  path = "HistonePTM_Abundance_PRMT5.xlsx"
)



```

```{r QC plots}

make_simple_qc <- function(df_long, value, ylab) {
  qc <- df_long %>%
    group_by(Histone, Treatment, Time, Rep) %>%
    summarize(val = mean({{value}}, na.rm=TRUE), .groups="drop") %>%
    mutate(
      Cond = factor(paste(Treatment, Time), levels = c("DMSO 9 hr", "DMSO 24 hr", "GSK591 9 hr", "GSK591 24 hr")),
      Histone = factor(Histone, levels = c("H3","H4"))
    )

  ggplot(qc, aes(x=Cond, y=val, fill=Treatment)) +
    stat_summary(fun=mean, geom="col", color="black", width=0.6) +
    stat_summary(fun.data=mean_se, geom="errorbar", width=0.2) +
    scale_fill_manual(values=c(DMSO="#CCCCCC", GSK591="forestgreen")) +
    facet_wrap(~Histone, scales="free_y") +
    labs(x=NULL, y=ylab) +
    theme_minimal(base_size=12) +
    theme(axis.text.x = element_text(angle=45, hjust=1), legend.position="none")
}

print(make_simple_qc(frac_df, fracHeavy, "Fraction heavy"))
print(make_simple_qc(abundance_long, Abundance, "Peptide abundance"))
print(make_simple_qc(abund_occ, occ, "Occupancy"))

 
    pdf(paste0(outdir, "/plots/", "QC_fraction_heavy_H3H4.pdf"), width = 4, height = 5)

    print(make_simple_qc(frac_df, fracHeavy, "Fraction heavy"))
    
    dev.off()


```

```{r peptide heatmaps}

# Compute summary of fraction heavy
frac_summary <- frac_df %>%
  group_by(Peptide_sequence, PTM, Histone, Treatment, Time) %>%
  dplyr::summarize(
    mean_fracH = mean(fracHeavy, na.rm = TRUE),
    sd_fracH   = sd(fracHeavy,   na.rm = TRUE),
    .groups    = "drop"
  ) %>%
  dplyr::mutate(RowID = glue("{Peptide_sequence} | {PTM}"))

# Heatmap function focused on Fraction Heavy
plot_fraction_heavy_heatmaps <- function(summary_df) {
  peptides <- unique(summary_df$Peptide_sequence)
  heat_colors <- colorRampPalette(rev(RColorBrewer::brewer.pal(8, "RdYlBu")))(50)

  for (pep in peptides) {
    df_sub <- summary_df %>% filter(Peptide_sequence == pep)

    if (nrow(df_sub) == 0) {
      message(glue("No data for peptide: {pep}"))
      next
    }

    ptm_order <- df_sub %>%
      dplyr::distinct(PTM) %>%
      dplyr::mutate(
        pos = if_else(str_to_lower(PTM) == "unmod", 0L,
                      as.integer(str_extract(PTM, "(?<=K)\\d+"))),
        n_sites = str_count(PTM, "K\\d+")
      ) %>%
      dplyr::arrange(pos, n_sites) %>%
      dplyr::pull(PTM)

    mat <- df_sub %>%
      dplyr::mutate(RowID = PTM, Cond = paste(Treatment, Time)) %>%
      dplyr::select(RowID, Cond, mean_fracH) %>%
      pivot_wider(names_from = Cond, values_from = mean_fracH) %>%
      column_to_rownames("RowID") %>%
      as.matrix()

    col_order <- c("DMSO 9 hr", "GSK591 9 hr","DMSO 24 hr", "GSK591 24 hr")
    mat <- mat[ptm_order, col_order[col_order %in% colnames(mat)], drop = FALSE]

    if (nrow(mat) == 0 || ncol(mat) == 0) next

    heatmap_plot <- pheatmap(
      mat,
      cluster_rows = FALSE,
      cluster_cols = FALSE,
      color = heat_colors,
      main = glue("{pep}: Fraction heavy"),
      na_col = "grey90",
      fontsize_row = 8,
      fontsize_col = 10,
      border_color = NA,
      gaps_col = 2
    )
    print(heatmap_plot)
    
    #can't use ggsave for pheatmap
    
    pdf(paste0(outdir, "/plots/", pep, "_fraction_heavy_heatmap.pdf"), width = 4, height = 5)

    print(heatmap_plot)
    dev.off()
    
}
}

# Run the function
plot_fraction_heavy_heatmaps(frac_summary)

# write csv file
  write_csv(frac_summary, file = file.path(paste0(outdir,"/csv"), "FractionHeavy_Heatmap.csv"))

```


```{r stacked bar chart old new PTMs}

#Function to create a stack bar chart of Old (light) and New (Heavy) histone PTMs

# Normalize within each Peptide + Rep + Channel group
ptm_frac <- int_long %>%
  filter(!is.na(Intensity)) %>%
  group_by(Peptide_sequence, PTM, Heavy_Light, Treatment, Time, Rep) %>%
  summarize(ptm_int = sum(Intensity, na.rm=TRUE), .groups = "drop") %>%
  group_by(Peptide_sequence, Heavy_Light, Treatment, Time, Rep) %>%
  mutate(
    total_int  = sum(ptm_int),
    frac_pct   = 100 * ptm_int / total_int
  ) %>%
  ungroup()

# summarize across replicates
ptm_frac_summary <- ptm_frac %>%
  group_by(Peptide_sequence, PTM, Heavy_Light, Treatment, Time) %>%
  summarize(
    mean_pct = mean(frac_pct, na.rm=TRUE),
    sd_pct   = sd(frac_pct, na.rm=TRUE),
    .groups = "drop"
  )

####### plotting function

plot_stacked_peptide <- function(peptide_name) {
  df_plot <- ptm_frac_summary %>%
    dplyr::filter(Peptide_sequence == peptide_name)

  if (nrow(df_plot) == 0) {
    message(glue::glue("No data for peptide: {peptide_name}"))
    return(NULL)
  }

  df_plot <- df_plot %>%
    dplyr::mutate(
      Age = ifelse(tolower(Heavy_Light) == "light", "Old", "New"),
      Cond_Age = paste(Treatment, Time, Age),
      Cond_Age = factor(
        Cond_Age,
        levels = c(
          "DMSO 9 hr Old",  "GSK591 9 hr Old",
          "DMSO 9 hr New",  "GSK591 9 hr New",
          "DMSO 24 hr Old", "GSK591 24 hr Old",
          "DMSO 24 hr New", "GSK591 24 hr New"
        )
      )
    )

  # Generate dynamic grayscale palette based on number of PTMs
  ptm_colors <- colorRampPalette(c( "gray85", "gray70", "gray40","black"))(length(unique(df_plot$PTM)))

  stacked_plot <- ggplot(df_plot, aes(x = Cond_Age, y = mean_pct, fill = PTM)) +
    geom_bar(stat = "identity", width = 0.6) +
    #geom_errorbar(
    #  aes(ymin = mean_pct - sd_pct, ymax = mean_pct + sd_pct),
    #  width = 0.2,
    #  position = "stack"
    #) +
    scale_fill_manual(values = ptm_colors) +
    labs(
      title = glue::glue("Peptide {peptide_name}"),
      y = "% of total peptide",
      x = NULL
    ) +
    theme_minimal(base_size = 14) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.title = element_blank()
    )

  print(stacked_plot)
  
  ggsave(
    filename = glue::glue("plots/{pep}_stacked_barplot.pdf"),
    plot = stacked_plot,
    width = 6, height = 5, dpi = 300
  )
  
 }

#Cycle through all peptides to plot them

for (pep in unique(ptm_frac_summary$Peptide_sequence)) {
  plot_stacked_peptide(pep)
}

 # output raw data
  write_csv(ptm_frac_summary, file = file.path(paste0(outdir,"csv"), "StackedBar_PTM_OldNew.csv"))

```
